### 2016-05-20 writing/latex/abstracteng.tex
Slides and fixes t bibliography

A permutation is an arrangement of \(n\) objects.
Permutation classes of classical permutation patterns
capture many interesting properties, such as
stack-sortability, and have links to many different combinatorial objects.
Mesh patterns are an extension of classical patterns that allow additional
restrictions to be placed on occurrences of the pattern.
Two mesh patterns are coincident if they are avoided by the same set of
permutations. We provide sufficient conditions for
coincidence among mesh patterns, whilst also avoiding a longer classical
pattern. These conditions, along with two special cases are used to completely
classify coincidence amongst families containing a mesh pattern of length 2
and a classical pattern of length 3. Two patterns are Wilf-equivalent if they
have the same number of avoiders at every length, we completely
Wilf-classifies mesh patterns of length 2 when avoiding the classical pattern 231.
% Finally we attempt to show some non-trivial Wilf-equivalences between avoiders
% of sets of the form \(231,m_1\) and \(231,m_2\)

>>>>

A permutation is an arrangement of \(n\) objects.
Permutation classes of classical permutation patterns
capture many interesting properties, such as
stack-sortability, and have links to many different combinatorial objects.
Mesh patterns are an extension of classical patterns that allow additional
restrictions to be placed on occurrences of the pattern.
Two mesh patterns are coincident if they are avoided by the same set of
permutations. We provide sufficient conditions for
coincidence among mesh patterns, whilst also avoiding a longer classical
pattern. These conditions, along with two special cases are used to completely
classify coincidence amongst families containing a mesh pattern of length 2
and a classical pattern of length 3. Two patterns are Wilf-equivalent if they
have the same number of avoiders at every length, we completely
Wilf-classify mesh patterns of length 2 when avoiding the classical pattern 231.
Finally we attempt to show some non-trivial Wilf-equivalences between avoiders
of sets of the form \(231,m_1\) and \(321,m_2\), as well as discussing possible
future work that could be derived from this work.

### 2016-05-20 writing/latex/introduction.tex
Slides and fixes t bibliography

\section{Mesh Patterns}
Mesh patterns were introduced by \textcite{journals/combinatorics/BrandenC11} to
capture explicit expansions for certain permutation statistics. They are a
natural extension of classical permutation patterns.
A \emph{mesh pattern} is a pair
\begin{equation*}
    p = (\pi,R)\text{ with } \tau \in \mathfrak{S}_k \text{ and } R \subseteq
    [0,k]\times [0,k].
\end{equation*}

@@

We define containment (denoted \(p\preceq \pi\)), and avoidance, of the pattern \(p\) in the permutation
\(\tau\) on mesh patterns analogously to classical containment, and avoidance,
of \(\pi\) in \(\tau\) with the additional restrictions on the relative
position of the occurrence of \(\pi\) in \(\tau\). These restrictions say that
no elements of \(\tau\) are allowed in the regions of the plot corresponding
to shaded boxes in the mesh. Formally defined by \textcite{journals/combinatorics/BrandenC11},
an \emph{occurrence} of \(p\) in \(\tau\) is a subset \(\omega\) of the plot of
\(\tau, G(\tau) = \{(i,\tau(i)\mid i\in[1,n]\}\) such that there are order-preserving
injections \(\alpha,\beta:[1,k]\mapsto[1,n]\) satisfying the following two conditions.

@@

In the past people have studied different classes
of permutations that can be described by mesh patterns. \textcite{babstein2000}
considered \emph{vincular} patterns (also known as \emph{generalised} or \emph{dashed} patterns),
those where two adjacent entries in the pattern must be adjacent in the permutation,
\ie \(R\) is a union of vertical strips.
\textcite{MR2652101} look at classes of pattern where both columns and rows
can be shaded, these are called \emph{bivincular} patterns.
\emph{Bruhat-restricted} patterns were studied by \textcite{MR2264071} in order
to establish necessary conditions for a Schubert variety to be Gorenstein.
Mesh patterns also encompass a subset of \emph{barred} patterns introduced by
\textcite{MR2716312}, those with only one barred letter.

>>>>

\section{Mesh Patterns}
When looking at a classical pattern, for example \(\perm{2,3,1}\), any occurrences can be arbitrarily
placed. However it may be interesting to consider occurrences where the elements
corresponding to the \(2\) and \(3\) are adjacent in the permutation. In the past
these sort of questions have lead to a variety of definitions.  \textcite{babstein2000}
considered \emph{vincular} patterns (also known as \emph{generalised} or \emph{dashed} patterns),
those where two adjacent entries in the pattern must be adjacent in the permutation,
\ie \(R\) is a union of vertical strips.
\textcite{MR2652101} look at classes of pattern where both columns and rows
can be shaded, these are called \emph{bivincular} patterns.
\emph{Bruhat-restricted} patterns were studied by \textcite{MR2264071} in order
to establish necessary conditions for a Schubert variety to be Gorenstein.
Mesh patterns also encompass a subset of \emph{barred} patterns introduced by
\textcite{MR2716312}, those with only one barred letter.

All of these definitions are subsumed under the definition of \emph{mesh patterns},
introduced by \textcite{journals/combinatorics/BrandenC11} to
capture explicit expansions for certain permutation statistics. They are a
natural extension of classical permutation patterns.
A \emph{mesh pattern} is a pair
\begin{equation*}
    p = (\tau,R)\text{ with } \tau \in \mathfrak{S}_k \text{ and } R \subseteq
    [0,k]\times [0,k].
\end{equation*}

@@

The example of an occurrence of \(\perm{2,3,1}\) with adjacent elements \(31\)
can be represented as a mesh pattern by
\begin{equation*}
  \pattern{}{2,3,1}{1/0,1/1,1/2,1/3}
\end{equation*}

We define containment (denoted \(p\preceq \pi\)), and avoidance, of the pattern \(p\) in the permutation
\(\tau\) on mesh patterns analogously to classical containment, and avoidance,
of \(\pi\) in \(\tau\) with the additional restrictions on the relative
position of the occurrence of \(\pi\) in \(\tau\). These restrictions say that
no elements of \(\tau\) are allowed in the regions of the plot corresponding
to shaded boxes in the mesh.

Formally defined by \textcite{journals/combinatorics/BrandenC11},
an \emph{occurrence} of \(p\) in \(\tau\) is a subset \(\omega\) of the plot of
\(\tau, G(\tau) = \{(i,\tau(i)\mid i\in[1,n]\}\) such that there are order-preserving
injections \(\alpha,\beta:[1,k]\mapsto[1,n]\) satisfying the following two conditions.

@@

### 2016-05-20 writing/latex/wilf.tex
Slides and fixes t bibliography

The patterns in the following set are coincident,
\begin{equation*}
    \left\{
        \begin{array}{c}
        \textpattern{}{1}{},
        \textpattern{}{1}{0/0},
        \textpattern{}{1}{1/0},
        \textpattern{}{1}{0/0,1/0},
        \textpattern{}{1}{0/1},\\
        \textpattern{}{1}{0/0,0/1},
        \textpattern{}{1}{1/1},
        \textpattern{}{1}{1/0,1/1},
        \textpattern{}{1}{0/1,1/1}
    \end{array}
    \right\}
\end{equation*}
due to the fact that every permutation, except the empty
permutation must contain an occurrence of all of these patterns.

@@

\begin{example}
    Given the permutation \(\pi = \perm{5,4,2,1,3,9,6,8,7}\) this corresponds to the
    partition \(\{\{5,4,2,1\},\{3\},\{9,6\},\{7,8\}\}\).
\end{example}

We call the least element in each block the \emph{block bottom}

>>>>

The patterns in the following set are coincident,
\begin{equation*}
    \left\{
        \begin{array}{c}
        \textpattern{}{1}{},
        \textpattern{}{1}{0/0},
        \textpattern{}{1}{1/0},
        \textpattern{}{1}{0/0,1/0},
        \textpattern{}{1}{0/1},\\
        \textpattern{}{1}{0/0,0/1},
        \textpattern{}{1}{1/1},
        \textpattern{}{1}{1/0,1/1},
        \textpattern{}{1}{0/1,1/1}
    \end{array}
    \right\}
\end{equation*}
due to the fact that every permutation, except the empty
permutation, must contain an occurrence of all of these patterns.

@@

\begin{example}
    Given the permutation \(\pi = \perm{5,4,2,1,3,9,6,8,7}\) this corresponds to the
    partition \(\{\{5,4,2,1\},\{3\},\{9,6\},\{8,7\}\}\).
\end{example}

We call the least element in each block the \emph{block bottom}.

### 2016-05-20 writing/presentation/coincidences.tex
Slides and fixes t bibliography

\section{Coincidence Classes}
\label{sec:Coincidence Classes}
\subsection{Coincidence}
\label{sub:Coincidence}
\begin{frame}{Coincidence}
  \begin{definition}
    Two mesh patterns are said to be \emph{coincident} if they avoid the
    same set of patterns at every length.
  \end{definition}
  \begin{block}{}
    Classical patterns can never be coincident.

@@

\end{tikzpicture}
    \caption{If the conditions of The Second Dominating rule are satisfied the box
    \((a-1,b)\) can be shaded.}
\end{figure}
  \end{example}
\end{frame}

@@

>>>>

\section{Coincidence Classes}
\label{sec:Coincidence Classes}
\subsection{Coincidence}
\label{sub:Coincidence}
\begin{frame}{Coincidence}
  \begin{definition}
    Two mesh patterns are said to be \emph{coincident} if they avoid the
    same set of permutations at every length.
  \end{definition}
  \begin{block}{}
    Classical patterns can never be coincident.

@@

\end{tikzpicture}
    \caption{If the conditions of The Second Dominating rule are satisfied the box
    \((a,b)\) can be shaded.}
\end{figure}
  \end{example}
\end{frame}

@@

\subsection{Special Cases}
\label{sub:Special Cases}
\begin{frame}{Special Cases}
  There are \(7\) coincidences of the form \(\av{\perm{3,2,1},m}\) that are not
  explained by the rules.
  \begin{example}
    \begin{equation*}
      m_1 = \pattern{}{1,2}{0/1,0/2,2/0,2/1} \text{ and } m_2 = \pattern{}{1,2}{0/1,0/2,1/1,2/0,2/1}
    \end{equation*}
    This coincidence is explained by mathematical induction on the number of
    points in region corresponding to the middle box.
  We call this number \(n\).
  \end{example}
\end{frame}
\begin{frame}
  \begin{example}
    \begin{description}
    \only<1-2>{\item [Base Case \((n=0)\):] The base case holds since we can freely shade the
        box if it contains no points.}
    \only<2>{\item [Inductive Hypothesis \((n=k)\):] Suppose that we can find an occurrence
        of the second pattern if we have an occurrence of the first with \(k\) points
        in the middle box.}
    \only<3>{\item [Inductive Step \((n=k+1)\)] Suppose that we have \((k+1)\) points in
        the middle box. Choose the bottom most point in the middle box, this gives the
        mesh pattern
     \begin{equation*}
        \begin{tikzpicture}[scale=0.5]
            \modpattern[5]{}{1,2,3}{0/1,0/2,0/3,1/1,2/1,3/0,3/1,3/2}
            \draw (2.5,0.5) node {\(X\)};
        \end{tikzpicture}
    \end{equation*}
        Now we need to consider the box labelled \(X\). If this box is empty then
        we have an occurrence of \(m_2\) and are done.}
        \only<4>{\item [Inductive Step \((n=k+1)\) (cont.)] If this box contains
        any points then we gain some extra shading on the mesh pattern
        due to the dominating pattern
     \begin{equation*}
        \begin{tikzpicture}[scale=0.5]
            \modpattern[5]{2,3}{1,2,3}{0/1,0/2,0/3,1/1,1/2,1/3,2/1,3/0,3/1,3/2}
            \filldraw (2.5,0.5) circle (4 pt);
        \end{tikzpicture}
    \end{equation*}
        The two highlighted points form an occurrence of \(m_1\) with \(k\)
        points in the middle box, and thus by the Inductive Hypothesis we
        are done.}
\end{description}
  \end{example}
\end{frame}

### 2016-05-20 writing/presentation/introduction.tex
More Presentation
Slides and fixes t bibliography

  More intuitively \textquote[Knuth]{A \emph{permutation of \(n\) objects} is
  an arrangement of \(n\) distinct objects in a row}.

  We write permutations in \emph{one-line notation}, writing the entries of
  the entries of the permutation in order \(\pi\)
  \begin{equation*}
    \pi = \perm{\pi(1),\pi(2),\dotso,\pi(n)}
  \end{equation*}
  \begin{example} The \(6\) permutations on \(\nrange{3}\) are
    \begin{equation*}
      123, 132, 213, 231, 312, 321
    \end{equation*}
  \end{example}
\end{frame}

>>>>

  More intuitively \textquote[Knuth~\cite{Knuth}]{A \emph{permutation of \(n\) objects} is
  an arrangement of \(n\) distinct objects in a row}.

  We write permutations in \emph{one-line notation}, writing the entries of
  the entries of the permutation in order
  \begin{equation*}
    \pi = \perm{\pi(1),\pi(2),\dotso,\pi(n)}
  \end{equation*}
  \begin{example} The \(6\) permutations on \(\nrange{3}\) are
    \begin{equation*}
      123, 132, 213, 231, 312, 321
    \end{equation*}
  \end{example}
\end{frame}
